---
title: "MLproject"
author: "Alla Aleksandrova"
date: '15 september 2019 г '
output:
  html_document: default
  pdf_document: default
editor_options:
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#Goal
The goal is to predict the manner in which people did the exercise.

#Summary

*Data Preparation*  
1. Download and load data sets,  
2. Preprocess data set: create partition and filter out the training and test data set,  
3. Find out the less useful predictors,  
4. Look into correlations.  
  
*Model Building*  
5. Fit model on three methods:  
a) Random Forest,  
b) Decision Tree,  
c) Generalized Boosted Model;  
6. Print accuracy of the built models.  
  
*Predict*  
7. Predict the 20 quiz results,  
8. Print out of sample error and percentage of out of sample error.  
  
  
#Data preparation results
```{r plot}
library(caret); library(knitr);library(randomForest)
library(rattle); library(rpart.plot);library(rpart)
library(gbm); library(corrplot); library(ggplot2)
library(colorspace); library(e1071)
#load("pml-testing.csv")
#load("pml-training.csv")
#load data
trainData <- read.csv("pml-training.csv"); dim(trainData)
testData <- read.csv("pml-testing.csv"); dim(testData)
set.seed(100)
#create sets
inTrain<-createDataPartition(y=trainData$classe, p=0.7, list=FALSE)
trainSet<- trainData[inTrain, ]
cvSet<- trainData[-inTrain, ]
#find variables near zero
NZV <- nearZeroVar(trainSet)
trainSet <- trainSet[, -NZV]
cvSet  <- cvSet[, -NZV]
dim(trainSet)
#remove variables that are mostly NA (mNA)
mostlyNA <- sapply(trainSet, function(x) mean(is.na(x))) > 0.95
trainSet <- trainSet[, mostlyNA==FALSE]
cvSet <- cvSet[, mostlyNA==FALSE]
#remove identification only variables (columns 1 to 5)
trainSet <- trainSet[, -(1:5)]
cvSet <- cvSet[, -(1:5)]

corMatrix <- cor(trainSet[, -54])
corrplot(corMatrix, order = "FPC", method = "color", type = "lower", 
         tl.cex = 0.8, tl.col = rgb(0, 0, 0))

```

#Model building results
Decision Tree Accuracy
```{r plot Random Forest Accuracy}
trControlRF <-trainControl(method = "cv", number = 3, verbose = FALSE)
modFitRF <-train(classe ~ ., data = trainSet, method = "rf", trControl = trControlRF)
modFitRF$finalModel

predictRF <- predict(modFitRF, newdata = cvSet)
confMatrixRF <- confusionMatrix(predictRF, cvSet$classe)

plot(confMatrixRF$table, col = confMatrixRF$byClass, 
     main = paste("Random Forest Accuracy =",
                  round(confMatrixRF$overall['Accuracy'], 4)))
```

Decision Tree Accuracy
```{r plot Decision Tree Accuracy}
modFitDT <- rpart(classe ~ ., data=trainSet, method="class")
fancyRpartPlot(modFitDT)

predictDT <- predict(modFitDT, newdata=cvSet, type="class")
confMatrixDT <- confusionMatrix(predictDT, cvSet$classe)
plot(confMatrixDT$table, col = confMatrixDT$byClass, 
     main = paste("Decision Tree Accuracy =",
                  round(confMatrixDT$overall['Accuracy'], 4)))
```
Generalized Boosted Model Accuracy
```{r plot Generalized Boosted Model Accuracy}
trControlGBM <- trainControl(method = "repeatedcv", number = 5, repeats = 1)
modFitGBM  <- train(classe ~ ., data=trainSet, method = "gbm",
                    trControl = trControlGBM, verbose = FALSE)

predictGBM <- predict(modFitGBM, newdata=cvSet)
confMatrixGBM <- confusionMatrix(predictGBM, cvSet$classe)
plot(confMatrixGBM$table, col = confMatrixGBM$byClass, 
     main = paste("Generalized Boosted Model Accuracy =",
                  round(confMatrixGBM$overall['Accuracy'], 4)))
```


After fitting our models the best accuracy is RandomForest Accuracy. The accuracy is 99.8%, thus my predicted accuracy for the out-of-sample error is 0.2%. I will use Random Forests to predict on the test set.

```{r summary predict}
#summary(predictTEST)
modFitRF <-train(classe ~ ., data = trainSet, method = "rf", trControl = trControlRF)
predictTEST <- predict(modFitRF, newdata=testData)
predictTEST <- c(as.character(predictTEST))
predictTEST
```

My out of sample error
```{r out of sample error}
outOfSampleError.accuracy <- sum(predictTEST == cvSet$classe)/length(predictTEST)
outOfSampleError <- 1 - outOfSampleError.accuracy
outOfSampleError
```